---
title: "Analyse_data_LIDAR"
author: "F.Boudon & B.Pallas  "
date: "March 2018"
colorlinks: true
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: TRUE
---

# Overview
```{r,  results='hide', message=FALSE, warning=FALSE}
library(PerformanceAnalytics)
library(lme4)
library(car)
library(FactoMineR)
library(rgl)
library('fBasics')
library(rglwidget)
library(agricolae)
library(minpack.lm)
library(matrixStats)
```


## data visualization
```{r}
setwd("D:/Benoit/2018/agreenium_training_2018/data")
data_lidar = read.csv("data_lidar.csv", header=T, sep=";")
str(data_lidar)
```

Definition of variables and units: </li>
<li> volume : volume of the convex hull  </li> 
<li> diameter : projected diameter on the soil of the tree canopy (m) </li>
<li> height : height of the trees (m) </li>
<li> estimated_leaf_area : estimated leaf area of the tree (m²) </li>
<li> STAR : silhouette to leaf area ratio (m²/m²) </li>
<li> max_excentricity : maximal normalized (depending on tree size) excentricity  </li>


##Experimental layout
```{r}
color = vector(mode="character", length=dim(data_lidar)[1])
color[data_lidar$treatment == "WD"] = "red"
color[data_lidar$treatment == "WW"] = "blue"
plot(data_lidar$col,data_lidar$row, ylim=c(0,120), col=color, main="Experimental_field", xlab="rank",ylab="line")
legend("topleft", legend = c("WS","WW"),text.col = c("red","blue") , cex = 0.8, bty = "n")
```

# Exploratory data analysis 
## Basic statistics
```{r}
general_statistics=as.data.frame(matrix(nrow=6, ncol=4))
colnames(general_statistics) = c("variable","mean","sd","Cv")
general_statistics["variable"] = colnames(data_lidar[,c(5: (dim(data_lidar)[2]))])
general_statistics["mean"] =  round(colMeans(data_lidar[,c(5: (dim(data_lidar)[2]))]),2)
general_statistics["sd"] = round(colSds(as.matrix(data_lidar[,c(5: (dim(data_lidar)[2]))]), na.rm=TRUE),2)
general_statistics["Cv"] = round(general_statistics["sd"] / general_statistics["mean"],2)
print(general_statistics)
```

## Correlation plot
```{r, message=FALSE, warning=FALSE}
chart.Correlation(data_lidar[,c(5: (dim(data_lidar)[2]))], histogram = TRUE, method = c("pearson"))
```

## Multivariate analysis
### PCA
```{r}
data_pca = data_lidar[,c(5: (dim(data_lidar)[2]))]
length = dim(data_pca)[2]
data_pca = na.omit(data_pca)

pca = PCA(data_pca, graph=FALSE)
plot(pca, choix="var", axes=c(1,2))
plot(pca, choix="var", axes=c(1,3))
```

### PCA with clustering
```{r, warning = FALSE}
nb_groups = 6 

mydata =scale(data_pca, center = TRUE, scale = TRUE)
id_names = paste0(data_lidar[,1],"-", data_lidar[,2], "-",data_lidar[,3])
rownames(mydata) = id_names
d=dist(mydata)
fit=hclust(d,method = "ward")
groups=cutree(fit,k=nb_groups)
data_pca$groups = groups
pca = PCA(data_pca, graph=FALSE)

plot(pca$ind$coord[,"Dim.1"], pca$ind$coord[,"Dim.2"], xlab="axe1", ylab="axe2",xlim=c(1.2*min(pca$ind$coord[,"Dim.1"]), 0.8*max(pca$ind$coord[,"Dim.1"])), ylim = c(1.2*min(pca$ind$coord[,"Dim.2"]), 1.2*max(pca$ind$coord[,"Dim.2"])), col=data_pca$groups)
abline(h=0,v=0, lty=2)
legend1 = c("groupe1")
for (i in (2:nb_groups)) {legend1 = c(legend1, paste("groupe",i, sep=""))}
legend("topleft", legend = legend1,text.col = c(1:nb_groups), cex = 0.8, bty = "n")

data_pca$groups = groups
mean_values = aggregate(data_pca[,1:length], list(data_pca$groups), mean)
print(mean_values)

```

# Basic genetics analysis (heritability)
```{r}
data_heritability = as.data.frame(matrix(nrow = 5, ncol = 2))
colnames(data_heritability) = c("variable","heritability")
list_variables = c("volume", "diameter", "height", "estimated_leaf_area", "STAR","max_excentricity")
acc = 0
 
for (i in list_variables)
{ acc = acc  + 1
  model0 <-lmer(data_lidar[,i] ~  (1|name) + treatment, data=data_lidar )
  variance_results = as.data.frame(VarCorr(model0))
  Var_G = variance_results$vcov[1]
  Var_R = variance_results$vcov[2]
  hG = Var_G / (Var_G + Var_R/4)
  hist(residuals(model0), , xlab = "residuals", main = i)
  data_heritability[acc,] = c(i,round(hG,2))
  
}
print(data_heritability)
```

# Analysis of the water treatment impact
```{r}
data_stress = as.data.frame(matrix(nrow = 5, ncol = 2))
colnames(data_stress) = c("variable","relative_effect")
list_variables = c("volume", "diameter", "height", "estimated_leaf_area", "STAR","max_excentricity")
acc = 0
for (i in list_variables)
{ acc = acc  + 1
  boxplot(data_lidar[,i] ~ treatment, data= data_lidar, xlab="treatment",ylab=i, main=i, col=c("red","blue"))
  model0 <-lmer(data_lidar[,i] ~  (1|name) + treatment, data=data_lidar )
  print(Anova(model0)["Pr(>Chisq)"])
  data_stress[acc,] = c(i, - round(fixef(model0)[2]/mean(data_lidar[,i]),2)) ### - because we estimate the effect of the WW treatment
}
print(data_stress)
```
